---
title: "Housing data analysis - Women Who Code workshop"
author: "Darya Vanichkina"
date: "13/03/2018"
output:
  html_document: default
  pdf_document: default
---

```{r setup,  echo=TRUE, results='hide',message=FALSE, warning=FALSE}
load.libraries <- c('tidyverse', 'forcats', 'corrplot', 'caret', 'Metrics', 'randomForest', 'xgboost', 'glmnet', 'car')
# note car only for Darya to use the demo dataset
install.lib <- load.libraries[!load.libraries %in% installed.packages()]
for(libs in install.lib) install.packages(libs, dependences = TRUE)
sapply(load.libraries, library, character = TRUE)
knitr::opts_chunk$set(echo = TRUE)
```

## Basics ---- 
```{r VeryBasics}
# Arithmetic operations: R is a calculator
1 + 1
10**2
TRUE + TRUE
```

Gotcha: R is not python (works only for numeric):

```{r}
# "1" + "1"
```
Variables:

```{r Variables}
x <- 1 # the R "way" to do it ...
y = 2
print(x)
print(y)
x
```

[More about why](https://csgillespie.wordpress.com/2010/11/16/assignment-operators-in-r-vs/)

Data types: vectors

- "character" (aka string), numeric and boolean 

```{r Vectors}
a <- c(1,2,5.3,6,-2,4) # numeric vector
b <- c("one","two","three") # character vector
c <- c(TRUE,TRUE,TRUE,FALSE,TRUE,FALSE) #logical vector
print(paste(c(class(a), class(b), class(c))))
```

Getting help and the combine function:

```{r HelpC,  echo=TRUE, results='hide'}
# getting help
?c
x <- c(1,2,3)
class(x)
y <- c(x, "2")
class(y)
```


## Data types: data frames (the main class for data analysis) --- 


#### Task 1
Load the data in from csv. 

```{r ReadData, echo=TRUE, results='hide'}
# read.csv("file")
myprestige <- Prestige
class(myprestige)
myprestige$job <- row.names(myprestige)
head(myprestige)

# train.csv
# test.csv
# Set the directory here (to where YOU are)
getwd()
setwd("~/Dropbox/Documents/06_Teaching/WWC/housing/")
# the above is my file path; you will need to change it to wherever wwc-2018 is on your filesystem
trainH <- read.csv("train.csv")
testH <- read.csv("test.csv")
```





#### Task 2

1. What features are there in the data? 
2. What are the dimensions of the data? 
3. What are the column headers? 

Use the summary() and str() functions to explore...


```{r WhatFeat,  echo=TRUE, results='hide'}
summary(myprestige)
str(myprestige)
table(myprestige$type, useNA = "always")
dim(myprestige)
# trainH
summary(trainH)
str(trainH)
dim(trainH)
```

## What does the distribution of sale price look like?

#### Task 3

1. Is the sale price (the variable we're interested in prediting) normally distributed? 
2. Plot a histogram of the distribution using ggplot2.
3. Find its mean, standard deviation


```{r PlotSalehistogram}
myprestige %>%
  ggplot(aes(x = income)) + geom_histogram() 

myprestige %>%
  ggplot(aes(x = income)) + geom_histogram() + theme_minimal()

myprestige %>%
  ggplot(aes(x = income)) + geom_histogram(bins = 100) + theme_minimal()

myprestige %>%
  ggplot(aes(x = income)) + geom_histogram(bins = 100, aes(y = ..density..)) + theme_minimal() + geom_density(col = "red")

myprestige %>%
  ggplot(aes(x = income)) + geom_histogram(bins = 100, aes(y = ..density..)) + theme_minimal() + geom_density(col = "red") + stat_function(fun = dnorm, color = "blue", args=list(mean = mean(myprestige$income), sd = sd(myprestige$income)))


# trainH

trainH %>%
  ggplot(aes(x = SalePrice)) + geom_histogram(bins = 100, aes(y = ..density..)) + theme_minimal() + geom_density(col = "red") + stat_function(fun = dnorm, color = "blue", args=list(mean = mean(trainH$SalePrice), sd = sd(trainH$SalePrice)))


```

#### Task 4

1.Plot a quantile-quantile plot (QQ plot) to "assess" normality. 

    Note: This plot compares the data we have (Sample Quantiles) with a theoretical sample coming from a normal distribution.  Each point (x, y) corresponds to one of the quantiles of the second distribution (x-coordinate, theoretical) plotted against the same quantile of the first distribution (y-coordinate, our data). Thus the line is a parametric curve with the parameter which is the number of the interval for the quantile.

```{r QQPlot}
qqnorm(myprestige$income)
qqline(myprestige$income)

qqnorm(trainH$SalePrice)
qqline(trainH$SalePrice)

```

A standard way of transforming the data to be better approximated by a normal distribution is by using the log-transform? 

#### Task 5
1. Carry out this transformation 
2. Use a histogram and QQ plot to see whether it works...


```{r LogTransform}
myprestige$incomeLog <- log(myprestige$income + 1)
qqnorm(myprestige$incomeLog) 
qqline(myprestige$incomeLog)


trainH$LogSalePrice <-log(trainH$SalePrice + 1)
qqnorm(trainH$LogSalePrice) 
qqline(trainH$LogSalePrice)
trainH$SalePrice <- NULL
myprestige$income <- NULL
```


## Missing data

#### Task 6

What happens if we only use complete data? How much data is missing?


Topics used here (but not explored):
[Subsetting data frames](https://swcarpentry.github.io/r-novice-gapminder/06-data-subsetting/)
[The apply family](https://nsaunders.wordpress.com/2010/08/20/a-brief-introduction-to-apply-in-r/) 

```{r whatsmissing}
dim(myprestige)
dim(myprestige[complete.cases(myprestige),  ])
dim(myprestige[!complete.cases(myprestige),  ])

colSums(sapply(myprestige, is.na)) 

colSums(sapply(myprestige, is.na)) [colSums((sapply(myprestige, is.na))) > 0]

# trainH
colSums(sapply(trainH, is.na)) [colSums((sapply(trainH, is.na))) > 0]


```


How do we impute the missing data?

Read the metadata file and see that many of the NAs should be recoded as None since these features are lacking in the house. 

#### Task 9
Recode the NA values that should be None using mutate() and fct_explicit_na(). 

```{r missingBasedOnDescription}
trainH <- trainH %>% mutate(PoolQC = fct_explicit_na(PoolQC, na_level = "None"))
```

#### Task 10

For the GarageYrBlt - set NA values using replace_na() to zero. 

```{r missingSetToZero}

trainH <- trainH %>% replace_na(list(GarageYrBlt = 0))

```

#### Task 11

For Lot frontage - set it to be the median for the neighborhood using group_by() and mutate().

```{r LotFrontageMedian}
trainH %>% group_by(Neighborhood) %>% summarise(lotFmed = median(LotFrontage, na.rm = TRUE))

trainH <- trainH %>% group_by(Neighborhood) %>% mutate(LotFrontage = ifelse(is.na(LotFrontage), median(LotFrontage, na.rm = TRUE), LotFrontage))

```


***


## Basic exploratory data analysis of training data

#### Task 13
1. How does the sale price depend on living  area: X1stFlrSF, X2ndFlrSF, TotalBsmtSF? (use a scatterplot to visualise this)
2. Create a variable TotalSqFt which is a combination of these 
3. Does it better predict the house price? (again, just using scatterplot at this point)

```{r SalePriceExplore}
trainH %>% ggplot(aes(x = X1stFlrSF, y = LogSalePrice)) +geom_point() + theme_minimal()

trainH %>% ggplot(aes(x = X2ndFlrSF, y = LogSalePrice)) +geom_point() + theme_minimal()


```

What variables are correlated with each other and with price? 

#### Task 20
1. Plot a correlation plot using corrplot() for all numeric variables and
2. ... those that show the top correlation with LogSalePrice.

```{r Corrplot}
trainHCnumeric <- trainH[, sapply(trainH, is.numeric)]

corrplot(cor(trainHCnumeric, use="everything"), method = "circle", type = "lower", sig.level = 0.01, insig= "blank")

correlationmatrix <- as.data.frame(cor(trainHCnumeric, use = "everything"))
correlationmatrix$name <- row.names(correlationmatrix)

correlationmatrix %>% select(LogSalePrice, name) %>% arrange(desc(LogSalePrice))

```

#### Task 21

Use the createDataPartition() function to separate the training data into a training and testing subset. Allocate 50% of the data to each class. Run set.seed(12) before this.


```{r MakeTestTrain}
set.seed(12)
partionD <- createDataPartition(y = trainH$LogSalePrice, p = 0.5, list = FALSE)
trainHtrain <- trainHCnumeric[partionD,]
trainHtest <- trainHCnumeric[-partionD,]

```


#### Task 22

Fit a linear model considering the "top 10"" correlated (top 9, ignore LogSalePrice for obvious reasons). Code the variables (column names) manually.


```{r lm}
lm_overallQ <- lm(LogSalePrice ~ OverallQual, data = trainHtrain)
lm_QualGrLiv <- lm(LogSalePrice ~ OverallQual + GrLivArea, data = trainHtrain)
summary(lm_overallQ)
summary(lm_QualGrLiv)
```

#### Task 23

1. Use predict() to predict house prices using our top10 model on the "test" portion of the training dataset. 
2. Use rmse to assess the root mean square error (our metric of accuracy). 

```{r HowdWeGo}
predition1 <- predict(lm_overallQ, trainHtest, type = "response")
predition2 <- predict(lm_QualGrLiv, trainHtest, type = "response")
rmse(trainHtest$LogSalePrice, predition1)
rmse(trainHtest$LogSalePrice, predition2)

```


#### Task 24

1. Use randomForest() to train a random forest model on all of the variables. 
2. Use predict() and rmse() to make the prediction and assess the accuracy respectively. 
3. Was a linear (on 9 features) or random forest model more accurate?

```{r RandomForest}
# randFor <- randomForest(LogSalePrice ~ ., data=trainHCtrain)
# # Predict using the test set
# prediction_rf <- predict(randFor, trainHCtest)
# trainHCtest$randFor <- prediction_rf
# # rmse?
# rmse(trainHCtest$LogSalePrice, trainHCtest$randFor)
```

#### Task 25

1. Use xgboost to predict house prices from numeric features of training dataset. 
2. Use xgb.plot.importance() to assess which variables are most important for predicting house prices. 


```{r NumericOnlyXGboost}
# trainHCtrainNum <- as(as.matrix(trainHCtrain[ , sapply(trainHCtrain, is.numeric)]), "sparseMatrix")
# trainHCtestNum <-  as(as.matrix(trainHCtest[ , sapply(trainHCtest, is.numeric)]), "sparseMatrix")
# 
# trainD <- xgb.DMatrix(data = trainHCtrainNum, label = trainHCtrainNum[,"LogSalePrice"])
# 
# #Cross validate the model
# cv.sparse <- xgb.cv(data = trainD,
#                     nrounds = 600,
#                     min_child_weight = 0,
#                     max_depth = 10,
#                     eta = 0.02,
#                     subsample = .7,
#                     colsample_bytree = .7,
#                     booster = "gbtree",
#                     eval_metric = "rmse",
#                     verbose = TRUE,
#                     print_every_n = 50,
#                     nfold = 4,
#                     nthread = 2,
#                     objective="reg:linear")
# 
# #Train the model
# #Choose the parameters for the model
# param <- list(colsample_bytree = .7,
#              subsample = .7,
#              booster = "gbtree",
#              max_depth = 10,
#              eta = 0.02,
#              eval_metric = "rmse",
#              objective="reg:linear")
# 
# 
# #Train the model using those parameters
# bstSparse <-
#   xgb.train(params = param,
#             data = trainD,
#             nrounds = 600,
#             watchlist = list(train = trainD),
#             verbose = TRUE,
#             print_every_n = 50,
#             nthread = 2)
# 
# testD <- xgb.DMatrix(data = trainHCtestNum)
# 
# prediction <- predict(bstSparse, testD) #Make the prediction based on the half of the training data set aside
# 
# #Put testing prediction and test dataset all together
# 
# prediction <- as.data.frame(as.matrix(prediction))
# colnames(prediction) <- "xgboost"
# trainHCtest$xgboost <- prediction$xgboost
# 
# 
# #Test with RMSE
# rmse(trainHCtest$LogSalePrice, trainHCtest$xgboost)
# 
# # Feature importance
# importance_matrix <- xgb.importance(dimnames(trainD)[[2]], model = bstSparse)
# xgb.plot.importance(importance_matrix[1:10])
```


#### Task 26

1.Use the glmnet library to train a ridge regression model. 
2. Is it more or less accurate than XGBoost? 

```{r RidgeReg}
# trainHCtrainNumMatrix <- as.matrix(trainHCtrain[ , sapply(trainHCtrain, is.numeric)])
# trainHCtestNumMatrix  <-  as.matrix(trainHCtest[ , sapply(trainHCtest, is.numeric)])
# # cross validation for glmnet
# glm.cv.ridge <- cv.glmnet(trainHCtrainNum[,c(1:38,40)], trainHCtrainNum[,"LogSalePrice"], alpha = 0)
# penalty.ridge <- glm.cv.ridge$lambda.min
# glm.ridge <- glmnet(x = trainHCtrainNum[,c(1:38,40)], y = trainHCtrainNum[,"LogSalePrice"], alpha = 0, lambda = penalty.ridge )
# y_pred.ridge <- as.numeric(predict(glm.ridge, trainHCtestNum[,c(1:38,40)] ))
# rmse(trainHCtest$LogSalePrice, y_pred.ridge)
```